---
title: "faces_exploring_data"
output: html_document
date: "2026-04-10"
---

#Note
We will be doing Weighed Bayesian (WBA) and Simple Bayesian (SBA) on face-trustworthiness (3 ratings)
Main difference is that weighted has 2 weights that affect the final trustworthiness vote: one for "weight of own first vote" (w1) and one for "weight of group vote influence" (w2)
For WBA these can vary, for SBA the evidence will be weighed the same regardless of where it comes from (from oneself or from social interaction)
Since people are social creatures, the SBA should be worse overall.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(
  tidyverse, # Data manipulation and plotting
  dplyr,
  here, # Robust file paths
  ggplot2 # Robust file paths
)
```

#some data exploration
```{r}
#there are 80 faces
#0-8 scores
#feedback is NA when the participant is the first one to see the face
#mby plot this to see better

df <- read_csv("data/cogsci_clean.csv")

df %>% 
  group_by(FaceID) %>%
  summarize(mean_TRUST_first = mean(FirstRating))

df %>% 
  group_by(FaceID) %>%
  summarize(mean_TRUST_second = mean(SecondRating))

df %>% 
  group_by(FaceID) %>%
  summarize(mean_TRUST_change = mean(FirstRating)-mean(SecondRating))
```

#generate data that fits WBA and data that fits SBA specifically
#data has to be generated differently for each model(?)
```{r}
#first data for WBA should have a set of random face scores (we can maybe just use the mean score from the original exp for each face?, with some randomness around them for individual variability)
#the "randomness" results in a differing group score (make sure there's enough variation)
#the WBA agent weighs the evidence differently: they really care about the social score (0.75:0.25 for example)
#so in the 2nd rating, they generally (with some randomness) gravitate towards the "Feedback" in their score

#feedback is NA when the participant is the first one to see the face!! It gets updated over time
```

```{r}
#2nd data for SBA should have a set of random face scores (we can maybe just use the mean score from the original exp for each face?, with some randomness around them for individual variability)
#the "randomness" results in a differing group score (make sure there's enough variation)
#the SBA agent weighs the evidence the same (1:1)
#so in the 2nd rating, they generally gravitate to the social score less?

#feedback is NA when the participant is the first one to see the face!! It gets updated over time
```

