---
title: "face_simulating_data_trying"
output: html_document
date: "2026-04-16"
---

#doing it correctly now, as a beta (trying)
hopefully this is correct-ish at least?

this produces the data for the 3 different agents we need for the experiment: selfish, balanced, social (defaults from Ricardo's notebook)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(
  tidyverse, # Data manipulation and plotting
  dplyr,
  here, # Robust file paths
  ggplot2 #ggplot
)
```

#!!!!!!
THIS DOESNT HAVE THE RHO & KAPPA FIX FOR WBA YET (need to make the betaFaceModel rely on a kappa and rho values, like in notebook)
--> this will change the logic & values for w_direct and w_social (the selfish, balanced, social agent specifications, since they can be defined with rho and kappa)
(just follow the notebook i think?)
#!!!!!!

#!!!
AFTERWARDS need to make the 2 models (WBA and SBA) and fit them to each of the 3 datasets separately (to figure out "which model generated it") + model recovery & then in the end fit them to the real cogsci data to see how they deal with it
#!!!

```{r}
# Core belief-update function for face trustworthiness
# same priors etc. as example
# treating the ratings as counts (hopefully its right? it still hasnt been made into ratios as per 9th lecture)
betaFaceModel <- function(alpha_prior = 0.5, beta_prior = 0.5,
                          trust1, total1,
                          group_trust_mean, n_total_rating = 7, 
                          rho, kappa) {
  
  # adding rho into the function & REMOVING W_SOCIAL AND W_DIRECT!!! (is this correct?)
  #effective_w_social <- w_social * rho
  #effective_w_direct <- w_direct * (1 - rho)
  
  # posterior update (structure taken basically from examples) plus implemented kappa 
  alpha_post <- alpha_prior + kappa * ((1 - rho) * trust1 +
                                      rho * group_trust_mean)
  beta_post  <- beta_prior  + kappa * ((1 - rho) * (total1 - trust1) +
                                      rho * (n_total_rating - group_trust_mean))



  expected_trust <- alpha_post / (alpha_post + beta_post)
  ci_lower       <- qbeta(0.025, alpha_post, beta_post)
  ci_upper       <- qbeta(0.975, alpha_post, beta_post)

  list(
    alpha_post     = alpha_post,
    beta_post      = beta_post,
    expected_trust = expected_trust,
    ci_lower       = ci_lower,
    ci_upper       = ci_upper
  )
}

#the 3 agents!!!!! (note, here they dont sum up to 1)
agents <- tibble(
  agent_type = c("Self-Focused", "Balanced", "Socially-Influenced"),
  rho        = c(0.2, 0.5, 0.8),  # directly encodes the balance
  kappa      = c(2.0, 2.0, 2.5)
)

#the actual script that combines the experimental design with the beta
simulate_face_data_beta <- function(face_id, true_mean_face,
                                     w_direct, w_social, rho, kappa, #added rho and kappa here
                                     n_participants = 1000,
                                     n_total_rating = 7) {
  
  first_ratings  <- numeric(n_participants)
  second_ratings <- numeric(n_participants)
  group_trust_means <- numeric(n_participants)
  feedback_vec   <- numeric(n_participants)
  
  for (p in 1:n_participants) {
    
    # FIRST RATING (own perceptual evidence)
    # (multiple steps to fix a NA issue)
    prob <- true_mean_face / n_total_rating
    prob <- max(0, min(1, prob))
    first <- rbinom(1, 7, prob)
    # print(first)
    
    # GROUP MEAN (social info)
    if (p == 1) {
      group_trust_mean <- first
      feedback   <- 0 #no feedback for first rating
    } else {
      group_trust_mean <- round(mean(first_ratings[1:(p-1)])) #mean rating of everyone but you... so far (running average) (order matters)
      group_trust_mean <- max(0, min(n_total_rating, group_trust_mean))
      feedback   <- round(group_trust_mean - first) #feedback is the difference
      feedback   <- max(-3, min(3, feedback)) #feedback can only be -3 to 3
    }
    
    # BETA UPDATE
    result <- betaFaceModel(
      trust1 = first,
      total1 = n_total_rating,
      group_trust_mean = group_trust_mean,
      n_total_rating = n_total_rating,
      rho = rho,
      kappa = kappa #added row and kappa
    )
    
    # probability matching
    theta_sample <- rbeta(1, result$alpha_post, result$beta_post)
    second <- rbinom(1, 7, theta_sample)

    # SAVE
    first_ratings[p]  <- first
    second_ratings[p] <- second
    group_trust_means[p] <- group_trust_mean
    feedback_vec[p]   <- feedback
  }
  
  tibble(
  FaceID        = face_id,
  ParticipantID = 1:n_participants,
  FirstRating   = first_ratings,
  SecondRating  = second_ratings,
  GroupTrustMean = round(group_trust_means, 2),
  Feedback       = feedback_vec
)
}
```

```{r}
generate_all_agents <- function(agents,
                                n_faces = 80,
                                n_participants = 1000) {
  
  # mean face trustowrthiness, latent (same across agents!)
  face_params <- tibble(
    FaceID = 1:n_faces,
    raw_face_value = rnorm(n_faces, mean = 4, sd = 1.2)
  ) %>%
    mutate(
      true_mean_face = pmin(pmax(raw_face_value, 1), 8), #to keep it in 1-8 (this does truncate tails though)
    )

  # loop over agent types
  agents %>%
    mutate(data = pmap(
      list(agent_type, rho, kappa),
      function(agent_type, rho, kappa) {
        
        df <- map_dfr(1:n_faces, function(f) {
          simulate_face_data_beta(
            face_id = f,
            true_mean_face = face_params$true_mean_face[f],
            rho = rho,
            kappa = kappa,
            n_participants = n_participants
          )
        })
        
        df %>%
          mutate(
            agent_type = agent_type,
            FirstRating  = FirstRating + 1, #so its not 0-7 but 1-8
            SecondRating = SecondRating + 1
          )}
    )) %>%
    select(agent_type, data)
}

all_data <- generate_all_agents(agents)

write_csv(all_data$data[[1]], "data/self_focused_WBA_agent.csv")
write_csv(all_data$data[[2]], "data/SBA_agent.csv")
write_csv(all_data$data[[3]], "data/socially_influenced_WBA_agent.csv")
```

#visualizing them (make the plots better, mostly just checking to make sure they arent the same)
```{r}
all_df <- bind_rows(all_data$data, .id = "agent_type") %>%
  mutate(agent_type = recode(agent_type,
                             "1" = "Self-Focused",
                             "2" = "Balanced",
                             "3" = "Socially-Influenced"))
```

```{r}
# overall summaries
all_data %>%
  mutate(summary = map(data, ~
    .x %>%
      summarize(
        mean_first  = mean(FirstRating),
        mean_second = mean(SecondRating)
      )
  )) %>%
  select(agent_type, summary) %>%
  unnest(summary)

# face-level changes
all_data %>%
  mutate(summary = map(data, ~
    .x %>%
      group_by(FaceID) %>%
      summarize(
        mean_first  = mean(FirstRating),
        mean_second = mean(SecondRating),
        mean_change = mean(SecondRating - FirstRating),
        .groups = "drop"
      )
  )) %>%
  select(agent_type, summary) %>%
  unnest(summary)

#for plotting
plot_data <- all_data %>%
  mutate(summary = map(data, ~
    .x %>%
      group_by(FaceID) %>%
      summarize(
        First  = mean(FirstRating),
        Second = mean(SecondRating),
        .groups = "drop"
      )
  )) %>%
  select(agent_type, summary) %>%
  unnest(summary) %>%
  pivot_longer(
    cols = c(First, Second),
    names_to = "Type",
    values_to = "Rating"
  )

```


#plotting overall ratings (1st, 2nd)
```{r}
ggplot(plot_data, aes(FaceID, Rating, color = interaction(agent_type, Type))) +
  geom_line(linewidth = 1) +
  geom_point(size = 1.5) +
  labs(
    title = "First vs Second Ratings (All Agent Types)",
    color = "Agent × Rating"
  ) +
  theme_minimal()+
  facet_wrap(~agent_type)
```
#first and 2nd rating (should maybe fix the labs to be x-axis 1-8 in steps of 1)
```{r}
bind_rows(all_data$data, .id = "agent_type") %>%
  pivot_longer(
    cols = c(FirstRating, SecondRating),
    names_to = "Type",
    values_to = "Rating"
  ) %>%
  ggplot(aes(Rating, fill = agent_type)) +
  geom_histogram(alpha = 0.4, bins = 20, position = "identity") +
  facet_grid(agent_type ~ Type) +
  labs(
    title = "Rating Distributions by Agent Type",
    fill = "Agent Type"
  ) +
  theme_minimal()
```

#plotting change from 1st to 2nd
```{r}
all_data %>%
  mutate(change = map(data, ~
    .x %>%
      group_by(FaceID) %>%
      summarize(
        mean_change = mean(SecondRating - FirstRating),
        .groups = "drop"
      )
  )) %>%
  select(agent_type, change) %>%
  unnest(change) %>%
  group_by(agent_type) %>%
  group_split() %>%
  purrr::walk(function(df) {

    overall_mean <- mean(df$mean_change)

    p <- df %>%
      ggplot(aes(
        x = reorder(FaceID, mean_change),
        y = mean_change,
        group = 1
      )) +
      geom_line() +
      geom_point(aes(
        color = ifelse(mean_change > 0,
                       "Increased Trust",
                       "Decreased Trust")
      ), size = 2.5) +
      geom_hline(
        yintercept = overall_mean,
        linetype = "dashed",
        linewidth = 0.7
      ) +
      scale_color_manual(values = c(
        "Increased Trust" = "#2E8B57",
        "Decreased Trust" = "#CD5C5C"
      )) +
      labs(
        title = paste("Mean Change in Trust Ratings:", unique(df$agent_type)),
        x = "Face ID",
        y = "Mean Change",
        color = "Trust Change"
      ) +
      theme_minimal() +
      theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()
      )
    print(p)
  })
  

#this does differ from the actual cogsci data (but i mean... that had a lot less samples?)
#also, were likely not capturing some psychological effect (like reacting strongly to negative disagreement: i.e. group thinks someone is much less trustworthy than you, so you adjust due to caring more about negative feedback... or smth)

#social people vote more extremely (look at max and min y-axis)
```


```{r}
range(all_df$GroupTrustMean)
range(all_df$FirstRating)
range(all_df$SecondRating)
```

