---
title: "faces_exploring_data"
output: html_document
date: "2026-04-10"
---

#Note
We will be doing Weighed Bayesian (WBA) and Simple Bayesian (SBA) on face-trustworthiness (2 rating sessions per participant, 0-8 scoring)
Main difference is that weighted has 2 weights that affect the final trustworthiness vote: one for "weight of own first vote" (w1) and one for "weight of group vote influence" (w2)
For WBA these can vary, for SBA the evidence will be weighed the same regardless of where it comes from (from oneself or from social interaction)
Since people are social creatures, the SBA should be worse overall.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(
  tidyverse, # Data manipulation and plotting
  dplyr,
  here, # Robust file paths
  ggplot2
)

#there are 80 faces
#0-8 scores
#-3 to +3 feedback
df <- read_csv("data/cogsci_clean.csv") %>%
  filter(!is.na(Feedback))

set.seed(123)
```

#some data exploration
```{r}
# Rating distr (first rating & second rating) per face
df %>% 
  group_by(FaceID) %>% 
  summarize(
    First = mean(FirstRating),
    Second = mean(SecondRating)
  ) %>% 
  ggplot(aes(FaceID)) +
  geom_point(aes(y = First, color = "First")) +
  geom_line(aes(y = First, color = "First"), linewidth = 1) +
  geom_point(aes(y = Second, color = "Second")) +
  geom_line(aes(y = Second, color = "Second"), linewidth = 1) +
  labs(y = "Mean Rating", color = NULL) +
  theme_minimal()
#they dont change very much

#histogram of each rating (roughly normal-ish? few extremes)
df %>%
  pivot_longer(c(FirstRating, SecondRating),
               names_to = "Type",
               values_to = "Rating") %>%
  ggplot(aes(Rating, fill = Type)) +
  geom_histogram(bins = 20, alpha = 0.7) +
  facet_wrap(~Type) +
  labs(
    title = "Distribution of First vs Second Ratings",
    x = "Rating",
    y = "Count",
    fill = NULL
  ) +
  theme_minimal()

#Mean change between 1st and 2nd (mostly trust lowers)
df %>% 
  group_by(FaceID) %>%
  summarize(mean_TRUST_change = mean(SecondRating) - mean(FirstRating)) %>%
  mutate(overall_mean = mean(mean_TRUST_change)) %>%
  ggplot(aes(x = reorder(FaceID, mean_TRUST_change), y = mean_TRUST_change, group = 1)) +
  geom_line() +
  geom_point(aes(color = ifelse(mean_TRUST_change > 0, "Increased Trust", "Decreased Trust")), 
             size = 2.5) +
  geom_hline(aes(yintercept = overall_mean), linetype = "dashed", color = "black", linewidth = 0.7) + #mean, add to legend later
  scale_color_manual(values = c("Increased Trust" = "#2E8B57", 
                               "Decreased Trust" = "#CD5C5C")) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  labs(
    title = "Mean Change in Trust Ratings (First - Second)",
    x = "Face ID",
    y = "Mean Change",
    color = "Trust Change"
  ) +
  theme_minimal()+ 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) #theyre kinda crowded and ugly looking so omitting the x lab

#to see which face is which (not really useful that though)
df %>%
  group_by(FaceID) %>%
  summarize(
    mean_first = mean(FirstRating),
    mean_second = mean(SecondRating),
    mean_change = round(mean_second - mean_first,2)
  ) %>%
  arrange(desc(mean_change))
```


```{r generating_wba_sba_data_func}
#generate a "mean trustworthiness fixation value" for each face
#for each face in df, generate the first score with some randomness around this fixation mean (face-specific)
#then, using the mean score of all participants thus far, give each participant FEEDBACK (the feedback is 0 if you're the first?)
#then, for each participant, do the trials again for their 2nd score: every participant receives feedback and should weigh it for their final scoring according to the WBA logic (0.25:0.75) of evidence. w1 is own choice, w2 is group mean (which is the "social" influence). this assumes that the participant is social and that the data-generation process is differently affected by differing evidence

simulate_face_trust_data <- function(face_id, mu, W1, W2) {

  first_ratings  <- numeric(n_participants)
  second_ratings <- numeric(n_participants)
  feedback_vec   <- numeric(n_participants)
  group_means    <- numeric(n_participants)

  for (p in 1:n_participants) {

    #FIRST RATING
    first <- rnorm(1, mu, 1.2)
    first <- round(first) #so its a full integer
    first <- max(0, min(8, first)) # so its 0-8

    #GROUP MEAN SO FAR
    if (p == 1) {
      group_mean <- first   #group mean is their rating, so feedback is 0?
      feedback   <- 0
    } else {
      group_mean <- mean(first_ratings[1:(p-1)]) #from participant 2 onwards, the group mean feedback is the mean of all previous participants' first ratings (seemed to vary similarly in the exp?)
      feedback   <- round(group_mean - first)
      feedback   <- max(-3, min(3, feedback))
    }

    #SECOND RATING WBA
    second <- W1 * first + W2 * group_mean + rnorm(1, 0, noise_sd)
    second <- round(second)
    second <- max(0, min(8, second)) 
    #cat(paste(first, "first", second, "second",  face_id, "face_id"), "\n") *#testing

    #SAVE VALUES
    first_ratings[p]  <- first
    second_ratings[p] <- second
    feedback_vec[p]   <- feedback
    group_means[p]    <- group_mean
  }
  
  #SAVE AS TIBBLE
  tibble(
    FaceID        = face_id,
    ParticipantID = 1:n_participants,
    FirstRating   = first_ratings,
    SecondRating  = second_ratings,
    Feedback      = feedback_vec,
    GroupMean     = round(group_means, 2)
  )
}

# latent "true" trustworthiness for each face (a random mean) (same for both WBA and SBA)
n_faces  <- 80 #to kinda match the og exp, change if needed

face_params <- tibble(
  FaceID = 1:n_faces,
  fixation_mean = rnorm(n_faces, mean = 4, sd = 1.2) %>% #to mimic the real data, which is kind of a "normal" distribution
  pmax(2) %>%
  pmin(7)
)
```


```{r WBA}
n_participants  <- 100
W1       <- 0.25 #weight of own choice
W2       <- 0.75 #weight of social influence
noise_sd <- 0.5

df_seq_WBA <- map_dfr(
  1:n_faces,
  ~ simulate_face_trust_data(.x, face_params$fixation_mean[.x], W1, W2)
)

write_csv(df_seq_WBA, "data/synthetic_WBA.csv")
```

```{r SBA}
#2nd data for SBA is the same but the evidence is weighed the same regardless of where it comes from (from oneself or from social interaction)
n_participants  <- 100
W1       <- 0.5 #weight of own choice
W2       <- 0.5 #weight of social influence
noise_sd <- 0.5

df_seq_SBA <- map_dfr(
  1:n_faces,
  ~ simulate_face_trust_data(.x, face_params$fixation_mean[.x], W1, W2)
)

write_csv(df_seq_SBA, "data/synthetic_SBA.csv")

```