---
title: "model_fitting"
output: html_document
date: "2026-04-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(
  tidyverse, # Data manipulation and plotting
  dplyr,
  here, # Robust file paths
  ggplot2, #ggplot
  cmdstanr,
  loo,
  bayesplot, #Plotting MCMC
  furrr,        # Parallel processing
  future        # Future backend
)

# Parallel Setup: Leave 2 cores free for system stability
n_cores <- parallel::detectCores() - 2
plan(multisession, workers = n_cores)

# Optional: Set a default ggplot theme
theme_set(theme_classic())
```

#now we fit the 2 models (WBA and SBA) and fit them to each dataset & then compare which model "best explains" which dataset (3 datasets: selfish, balanced, social)

#trying to get this shit to work
#note: SBA is just a WBA with specific parameters (rho and kappa), but it does exist as a separate model as well

#not doing expansive param recovery etc. due to time constraints (just the bare minimum)

```{r}
regenerate_simulations <-  FALSE #so it doesnt repeat what already exists

make_stan_data <- function(df) {
  list(
    n                 = nrow(df),
    trust1            = as.integer(df$FirstRating - 1L), #just to fix it to the right space (its adjusted for plots in the simulating data, 1-8 originally)
    trust2            = as.integer(df$SecondRating - 1L),
    group_trust_mean = as.integer(round(df$GroupTrustMean)),
    n_total_rating    = 7L,
    prior_kappa_mu    = 2.0,
    prior_kappa_sigma = 1.0
  )
}

datasets <- list(
  self_focused        = read_csv("data/self_focused_WBA_agent.csv"),
  balanced            = read_csv("data/SBA_agent.csv"),
  socially_influenced = read_csv("data/socially_influenced_WBA_agent.csv")
)

sba_model <- cmdstan_model("models/SBA_model2.stan")
wba_model <- cmdstan_model("models/WBA_model2.stan")

results <- map(names(datasets), function(dataset_name) { #does this for each dataset
  
  df <- datasets[[dataset_name]]
  d <- make_stan_data(df)
  d_sba <- d[!names(d) %in% c("prior_kappa_mu", "prior_kappa_sigma")]
  
  # filenames
  sba_file <- paste0("models/", dataset_name, "_sba_fit.rds")
  wba_file <- paste0("models/", dataset_name, "_wba_fit.rds")
  
  # SBA
  if (regenerate_simulations ||!file.exists(sba_file)) {
    sba_fit <- sba_model$sample(
      data          = d_sba, #to omit the data it doesnt need
      fixed_param   = TRUE, #no freefloating params
      chains        = 1,
      iter_warmup   = 0,
      iter_sampling = 500, #trying less iter for SBA (maybe needs less, maybe should do more)
      refresh       = 100
    )
    saveRDS(sba_fit, sba_file)
    cat("SBA fit completed and saved to:", sba_file, "\n")
  } else {
    sba_fit <- readRDS(sba_file)
    cat("Loaded existing SBA fit from:", sba_file, "\n")
  }
  
  # WBA
  if (regenerate_simulations || !file.exists(wba_file)) {
    wba_fit <- wba_model$sample(
      data            = d,
      adapt_delta     = 0.95,
      chains          = 4,
      parallel_chains = 4,
      iter_warmup     = 500,
      iter_sampling   = 500, #this takes a while (TRYING with less)
      refresh         = 100
    )
    saveRDS(wba_fit, wba_file)
    cat("WBA fit completed and saved to:", wba_file, "\n")
  } else {
    wba_fit <- readRDS(wba_file)
    cat("Loaded existing WBA fit from:", wba_file, "\n")
  }
  
  list(
    sba = sba_fit,
    wba = wba_fit
  )
})

names(results) <- names(datasets)

# LOO comparison per dataset (hopefully correct)
loo_results <- map(results, function(r) {
  
  log_lik_wba <- r$wba$draws("log_lik", format = "matrix")
  log_lik_sba <- r$sba$draws("log_lik", format = "matrix")
  
  if (ncol(log_lik_wba) > max_obs) {
    keep <- sample(seq_len(ncol(log_lik_wba)), max_obs)
    log_lik_wba <- log_lik_wba[, keep]
    log_lik_sba <- log_lik_sba[, keep]
  }
  
  # make sure values are valid
  stopifnot(all(is.finite(log_lik_wba)))
  stopifnot(all(is.finite(log_lik_sba)))
  
  # compute loo separately
  loo_sba <- loo(log_lik_sba)
  loo_wba <- loo(log_lik_wba)
  
  loo_compare(loo_sba, loo_wba)
})
```



```{r}
#quick so can get this out on time (no indepth param recovery on this, just checking the sampling fit summary)
check_sampling <- function(fit, model_name, dataset_name) {
  cat("\n======", model_name, "on", dataset_name, "======\n")
  
  print(fit$summary())
  fit$diagnostic_summary()  # prints divergences, treedepth, ebfmi automatically
  
  # trace plots (skip SBA since no free params)
  if (!is.null(tryCatch(fit$draws(c("rho", "kappa")), error = function(e) NULL))) {
    print(mcmc_trace(fit$draws(c("rho", "kappa"))) +
      ggtitle(paste(model_name, "-", dataset_name)))
  }
}
```

