---
title: "model_fitting"
output: html_document
date: "2026-04-28"
---
#now we fit the 2 models (WBA and SBA) to each dataset & then compare which model "best explains" which dataset (3 datasets: selfish, balanced, social)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(
  tidyverse, # Data manipulation and plotting
  dplyr,
  here, # Robust file paths
  ggplot2, #ggplot
  cmdstanr,
  loo,
  bayesplot #Plotting MCMC
)

# Parallel Setup: Leave 2 cores free for system stability
n_cores <- parallel::detectCores() - 2

# Optional: Set a default ggplot theme
theme_set(theme_classic())
```

#not doing expansive param recovery etc. due to time constraints (just the bare minimum)
# (also param recovery would probably be hampered by the model & data logic differing on the group mean trustworthiness value? -- fixed mean vs adaptive mean. can't really bother to fix that right now, could be an issue technically?)

#note: SBA is just a WBA with specific parameters (rho and kappa), but it does exist as a separate model as well

#the files are uploaded separately (google drive for now) as theyre quite large (gigs): too lazy to set up gitLFS right now

```{r}
#to force it to overwrite (regen) simulations
regenerate_simulations_SBA <-  FALSE 
regenerate_simulations_WBA <-  FALSE

#1. SETTING THE STAN DATA
make_stan_data <- function(df) {
  list(
    n                 = nrow(df),
    trust1            = as.integer(df$FirstRating - 1L), #just to fix it to the right space (its adjusted for plots in the simulating data, 1-8 originally) --> recheck if this works ok (it should be fine?)
    trust2            = as.integer(df$SecondRating - 1L),
    group_trust_mean = as.integer(round(df$GroupTrustMean)), #this should be correct and not need -1L (recheck later though)
    n_total_rating    = 7L,
    prior_kappa_mu    = 2.0,
    prior_kappa_sigma = 1.0
  ) #am i doing this correctly? are these all the priors we should set, or? (alpha and beta priors are hardcoded 0.5 in the models)
}

datasets <- list(
  self_focused        = read_csv(here("data/self_focused_WBA_agent.csv")),
  balanced            = read_csv(here("data/SBA_agent.csv")),
  socially_influenced = read_csv(here("data/socially_influenced_WBA_agent.csv"))
)

sba_model <- cmdstan_model(here("models/SBA_model2.stan"))
wba_model <- cmdstan_model(here("models/WBA_model2.stan"))

dir.create(here("models/fits"), recursive = TRUE, showWarnings = FALSE)

#2. FITTING MODELS
results <- map(names(datasets), function(dataset_name) {
  
  df <- datasets[[dataset_name]]
  d <- make_stan_data(df)
  d_sba <- d[!names(d) %in% c("prior_kappa_mu", "prior_kappa_sigma")] #doesnt use any of these in the model
  
  # filenames
  sba_file <- here("models/fits", paste0(dataset_name, "_sba_fit.rds"))
  wba_file <- here("models/fits", paste0(dataset_name, "_wba_fit.rds"))
  
  # SBA (though its effectively a type of WBA)
  if (regenerate_simulations_SBA ||!file.exists(sba_file)) {
    sba_fit <- sba_model$sample(
      data          = d_sba, #to omit the data it doesnt need
      fixed_param   = TRUE, #no freefloating params (so no warmup either)
      chains        = 2,
      parallel_chains = min(2, n_cores),  #low to save time
      iter_sampling = 1000,
      refresh       = 100,
      seed = 123
      )
    t1 <- Sys.time()
    #saveRDS(sba_fit, sba_file) #quicker, but flushes the temp files when you close R
    sba_fit$save_object(sba_file) #slow but thorough (& files take up a lot of space)
    cat("WBA fit saved to:", sba_file, "\n")
    cat("Save time:", difftime(Sys.time(), t1, units = "secs"), "s\n")
  } else {
    sba_fit <- readRDS(sba_file)
    cat("Loaded existing SBA fit from:", sba_file, "\n")
  }
  
  # WBA
  if (regenerate_simulations_WBA || !file.exists(wba_file)) {
    wba_fit <- wba_model$sample(
      data            = d,
      chains          = 2, #to save time 
      parallel_chains = min(2, n_cores), #low to save time
      iter_warmup     = 500, #warmups are slightly lower than 50/50 but still takes forever...
      iter_sampling   = 1000, #this takes a while, should have done less (could be a model issue?)
      refresh         = 100,
      seed = 123
      )
    t1 <- Sys.time()
    #saveRDS(wba_fit, wba_file) #is flushed once R is closed
    wba_fit$save_object(wba_file) 
    cat("WBA fit saved to:", wba_file, "\n")
    cat("Save time:", difftime(Sys.time(), t1, units = "secs"), "s\n")
  } else {
    wba_fit <- readRDS(wba_file)
    cat("Loaded existing WBA fit from:", wba_file, "\n")
  }
  
  list(
  sba_dir = sba_dir,
  wba_dir = wba_dir
)
  
})
names(results) <- names(datasets)
```


```{r}
#3. CHECKING DIAGNOSTICS (WBA)
check_diagnostics <- function(fit, label) {
  diag_smry <- fit$diagnostic_summary()
  min_ebfmi <- suppressWarnings(min(diag_smry$ebfmi, na.rm = TRUE))

  smry <- fit$summary() |> filter(!is.na(rhat))
  n_fail <- smry |> filter(rhat > 1.01 | ess_bulk < 400 | ess_tail < 400) |> nrow()

  cat(sprintf("%-25s | Divs: %3d | Min E-BFMI: %.2f | Failures: %d\n",
              label,
              sum(diag_smry$num_divergent),
              min_ebfmi,
              n_fail))
}

cat("\n=== WBA Diagnostics ===\n")
imap(results, function(r, name) {
  check_diagnostics(r$wba, paste0("WBA_", name))
})

#4. LOO COMPARISON PER DATASET (hopefully correct)
loo_results <- imap(results, function(r, dataset_name) {

  log_lik_wba <- r$wba$draws("log_lik", format = "matrix")
  log_lik_sba <- r$sba$draws("log_lik", format = "matrix")

  stopifnot(ncol(log_lik_wba) == ncol(log_lik_sba))
  stopifnot(all(is.finite(log_lik_wba)))
  stopifnot(all(is.finite(log_lik_sba)))

  loo_wba <- loo(log_lik_wba)
  loo_sba <- loo(log_lik_sba)

  comparison <- loo_compare(list(WBA = loo_wba, SBA = loo_sba))
  cat("\n=== LOO Comparison:", dataset_name, "===\n")
  print(comparison)
  comparison
})
```



```{r}
#quick so can get this out on time (no indepth param recovery on this, just checking the sampling fit summary)
check_sampling <- function(fit, model_name, dataset_name) {
  cat("\n======", model_name, "on", dataset_name, "======\n")
  
  print(fit$summary())
  print(fit$diagnostic_summary()) # prints divergences, treedepth, ebfmi
  
  # trace plots (skip SBA since no free params)
  if (!is.null(tryCatch(fit$draws(c("rho", "kappa")), error = function(e) NULL))) {
    print(mcmc_trace(fit$draws(c("rho", "kappa"))) +
      ggtitle(paste(model_name, "-", dataset_name)))
  }
}

imap(results, function(r, name) {
  check_sampling(r$wba, "WBA", name)
  check_sampling(r$sba, "SBA", name)
})
```

