---
title: "model_diagnostics_simple"
output: html_document
date: "2026-04-28"
---
#plotting the simple diagnostics
#not doing expansive param recovery etc. due to time constraints (just the bare minimum)
# (also param recovery would probably be hampered by the model & data logic differing on the group mean trustworthiness value? -- fixed mean vs adaptive mean. can't really bother to fix that right now, could be an issue technically?)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(
  tidyverse, # Data manipulation and plotting
  dplyr,
  here, # Robust file paths
  ggplot2, #ggplot
  cmdstanr,
  loo,
  bayesplot #Plotting MCMC
)
#paths of the fits (shouldve been a txt file...)
results <- readRDS(("models/fits/results_paths.rds"))
print(results)

#make it your path so you don't have to rerun things, etc. etc.
#this was a VERY stupid way of doing this!!
base_path <- "C:/Users/Igiba/Documents/sem_7_acm/Assignment_3_Bestest_study_group_ever_AdvCogMod/models/fits/"


#there was a typo here!!
results$self_focused <- list(
  sba_file = paste0(base_path, "self_focused_sba_fit.rds"),
  wba_file = paste0(base_path, "self_focused_wba_fit.rds")
)

results$balanced <- list(
  sba_file = paste0(base_path, "balanced_sba_fit.rds"),
  wba_file = paste0(base_path, "balanced_wba_fit.rds")
)

results$socially_influenced <- list(
  sba_file = paste0(base_path, "socially_influenced_sba_fit.rds"),
  wba_file = paste0(base_path, "socially_influenced_wba_fit.rds")
)
```

```{r}
#3. CHECKING DIAGNOSTICS (WBA)
check_diagnostics <- function(fit_slim, label) {
  diag_smry <- fit_slim$diagnostic_summary        # already computed at save time, no method call needed
  min_ebfmi <- suppressWarnings(min(diag_smry$ebfmi, na.rm = TRUE))
  smry <- fit_slim$summary |> filter(!is.na(rhat))
  n_fail <- smry |> filter(rhat > 1.01 | ess_bulk < 400 | ess_tail < 400) |> nrow()
  cat(sprintf("%-25s | Divs: %3d | Min E-BFMI: %.2f | Failures: %d\n",
              label,
              sum(diag_smry$num_divergent),
              min_ebfmi,
              n_fail))
}

cat("\n=== WBA Diagnostics ===\n")
invisible(
  imap(results, function(r, name) {
    wba_slim <- readRDS(r$wba_file)
    check_diagnostics(wba_slim, paste0("WBA_", name))
  })
)

#4. LOO COMPARISON PER DATASET (hopefully correct)
loo_results <- imap(results, function(r, dataset_name) {
  
  wba_slim <- readRDS(r$wba_file)
  sba_slim <- readRDS(r$sba_file)

  # extract loo objects
  wba_loo <- wba_slim$loo
  sba_loo <- sba_slim$loo

  # --- SAFETY CHECK: Pareto-k diagnostics ---
  check_k <- function(loo_obj, name) {
    k <- loo_obj$diagnostics$pareto_k
    cat("\n", name, "Pareto-k summary:\n")
    print(summary(k))
    
    if (any(k > 0.7, na.rm = TRUE)) {
      warning(name, ": high Pareto-k detected → LOO may be unreliable")
    }
  }

  cat("\n==============================\n")
  cat("DATASET:", dataset_name, "\n")

  check_k(wba_loo, "WBA")
  check_k(sba_loo, "SBA")

  # LOO comparison (safe)
  comparison <- loo_compare(list(WBA = wba_loo,
                                SBA = sba_loo))

  cat("\n=== LOO Comparison ===\n")
  print(comparison)

  # --- WEIGHTS (ONLY if safe) ---
  w <- NULL

  if (all(wba_loo$diagnostics$pareto_k < 0.7, na.rm = TRUE) &&
      all(sba_loo$diagnostics$pareto_k < 0.7, na.rm = TRUE)) {

    w <- loo_model_weights(list(WBA = wba_loo,
                                SBA = sba_loo),
                           method = "stacking")

    cat("\n=== Model Weights (stacking) ===\n")
    print(w$weights)

  } else {
    cat("\n⚠ Skipping weights: unstable PSIS (high Pareto-k)\n")
  }

  list(
    comparison = comparison,
    weights = w
  )
})

#struggling with getting the weights unfortunately (for SBA) as its pareto k blows the fuck up
#maybe a mistake somewhere in the process? SBA is fully deterministic (fixed=TRUE), or in the model specification?
#we can assume itd weigh nearly = 1 WBA for all of these, except maybe balanced (this should be higher SBA due to simplicity)

#this is worse now? For SBA, it sucks even more than it used to(?)
```

#trying to figure out why SBA-mimic has issues
#BALANCED
```{r}
sba1 <- readRDS(results$balanced$sba_file)
pointwise_ll <- sba1$loo$pointwise[, "elpd_loo"]

tibble(obs = seq_along(pointwise_ll), ll = pointwise_ll) |>
  ggplot(aes(x = obs, y = ll)) +
  geom_point(alpha = 0.4) +
  geom_hline(yintercept = -4, color = "red", linetype = "dashed") +
  labs(title = "SBA pointwise elpd_loo", y = "elpd per obs", x = "observation")

#most are fine? between 0 and -4
```
# BALANCED (most SBA-agreeing)
```{r}
df_balanced <- read_csv("data/SBA_agent.csv", n_max = 25000)

#clamping it the same
df_balanced <- df_balanced |>
  mutate(group_trust_mean_clamped = pmax(1, as.integer(round(GroupTrustMean))))

# joining pointwise ll to data
df_diag <- df_balanced |>
  mutate(
    obs         = row_number(),
    elpd        = pointwise_ll,
    bad         = elpd < -4,
    implied_theta = (0.5 + group_trust_mean_clamped + FirstRating) /
                    (1 + 2*7 + group_trust_mean_clamped + FirstRating +
                     (7 - group_trust_mean_clamped) + (7 - FirstRating)),
    signal_conflict = abs(FirstRating - group_trust_mean_clamped)
  )

# 1. where do bad observations sit in trust1 x group_mean space?
df_diag |>
  ggplot(aes(x = FirstRating, y = group_trust_mean_clamped, color = elpd)) +
  geom_jitter(alpha = 0.3, width = 0.2, height = 0.2) +
  scale_color_viridis_c() +
  labs(title = "elpd by trust1!!! and group mean")

# 2. does signal conflict predict bad fit?
df_diag |>
  ggplot(aes(x = signal_conflict, y = elpd)) +
  geom_jitter(alpha = 0.2, width = 0.1) +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "signal conflict vs elpd")

# 3. what does trust2 look like for bad vs good observations?
df_diag |>
  ggplot(aes(x = SecondRating, fill = bad)) +
  geom_histogram(position = "dodge", bins = 8) +
  labs(title = "trust2!!! distribution: bad vs good fit observations")

# 4. print the worst cases
df_diag |>
  filter(bad) |>
  arrange(elpd) |>
  select(obs, FirstRating, group_trust_mean_clamped, SecondRating, elpd, signal_conflict) |>
  head(20)

#issues throughout trust1, at big values of trust2 especially but also throughout
#signal conflict (plot 2) is not the reason for issues
```

#SELF-FOCUSED
```{r}
#load the elpd
sba2 <- readRDS(results$self_focused$sba_file)
pointwise_ll <- sba2$loo$pointwise[, "elpd_loo"]

df_selffocused <- read_csv("data/self_focused_WBA_agent.csv", n_max = 25000)

#clamping it the same
df_selffocused <- df_selffocused |>
  mutate(group_trust_mean_clamped = pmax(1, as.integer(round(GroupTrustMean))))

# joining pointwise ll to data
df_diag <- df_selffocused |>
  mutate(
    obs         = row_number(),
    elpd        = pointwise_ll,
    bad         = elpd < -4,
    implied_theta = (0.5 + group_trust_mean_clamped + FirstRating) /
                    (1 + 2*7 + group_trust_mean_clamped + FirstRating +
                     (7 - group_trust_mean_clamped) + (7 - FirstRating)),
    signal_conflict = abs(FirstRating - group_trust_mean_clamped)
  )

# 1. where do bad observations sit in trust1 x group_mean space?
df_diag |>
  ggplot(aes(x = FirstRating, y = group_trust_mean_clamped, color = elpd)) +
  geom_jitter(alpha = 0.3, width = 0.2, height = 0.2) +
  scale_color_viridis_c() +
  labs(title = "elpd by trust1!!! and group mean")

# 2. does signal conflict predict bad fit?
df_diag |>
  ggplot(aes(x = signal_conflict, y = elpd)) +
  geom_jitter(alpha = 0.2, width = 0.1) +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "signal conflict vs elpd")

# 3. what does trust2 look like for bad vs good observations?
df_diag |>
  ggplot(aes(x = SecondRating, fill = bad)) +
  geom_histogram(position = "dodge", bins = 8) +
  labs(title = "trust2!!! distribution: bad vs good fit observations")

# 4. print the worst cases
df_diag |>
  filter(bad) |>
  arrange(elpd) |>
  select(obs, FirstRating, group_trust_mean_clamped, SecondRating, elpd, signal_conflict) |>
  head(20)

#issues throughout trust1, at big values of trust2 especially but also throughout
#signal conflict (plot 2) is not the reason for issues
```

#SOCIALLY-INFLUENCED
```{r}
sba3 <- readRDS(results$socially_influenced$sba_file)
pointwise_ll <- sba3$loo$pointwise[, "elpd_loo"]

df_socialf <- read_csv("data/socially_influenced_WBA_agent.csv", n_max = 25000)

#clamping it the same
df_socialf <- df_socialf |>
  mutate(group_trust_mean_clamped = pmax(1, as.integer(round(GroupTrustMean))))

# joining pointwise ll to data (simulation data to model)
df_diag <- df_socialf |>
  mutate(
    obs         = row_number(),
    elpd        = pointwise_ll,
    bad         = elpd < -4,
    implied_theta = (0.5 + group_trust_mean_clamped + FirstRating) /
                    (1 + 2*7 + group_trust_mean_clamped + FirstRating +
                     (7 - group_trust_mean_clamped) + (7 - FirstRating)),
    signal_conflict = abs(FirstRating - group_trust_mean_clamped)
  )

# 1. where do bad observations sit in trust1 x group_mean space?
df_diag |>
  ggplot(aes(x = FirstRating, y = group_trust_mean_clamped, color = elpd)) +
  geom_jitter(alpha = 0.3, width = 0.2, height = 0.2) +
  scale_color_viridis_c() +
  labs(title = "elpd by trust1!!! and group mean")

# 2. does signal conflict predict bad fit?
df_diag |>
  ggplot(aes(x = signal_conflict, y = elpd)) +
  geom_jitter(alpha = 0.2, width = 0.1) +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "signal conflict vs elpd")

# 3. what does trust2 look like for bad vs good observations?
df_diag |>
  ggplot(aes(x = SecondRating, fill = bad)) +
  geom_histogram(position = "dodge", bins = 8) +
  labs(title = "trust2!!! distribution: bad vs good fit observations")

# 4. print the worst cases
df_diag |>
  filter(bad) |>
  arrange(elpd) |>
  select(obs, FirstRating, group_trust_mean_clamped, SecondRating, elpd, signal_conflict) |>
  head(20)

#issues throughout trust1, at big values of trust2 especially but also throughout
#signal conflict (plot 2) is not the reason for issues
```
#Note: Our SBA is a Fixed prior model with no sampled parameters, equally weighed evidence as per SBA, but also same kappa
#Every MCMC draw has identical alpha_post and beta_post values: there is no posterior uncertainty
# --> log_lik has zero variance across chains/draws for each observation
# (Warning: Can't fit generalized Pareto distribution because all tail values are the same.)



